---
title: "<center>**... in R**</center>"
author: "<center>Wyclife Agumba Oluoch (wyclifeoluoch@gmail.com) </center>"
date: "<center>`r Sys.time()`</center>"
bibliography: 
  - bib/packages.bib
nocite: '@*'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: darkly
    #highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
packages <- c("base",
              'knitr',
              'rmarkdown',
              'prettydoc',
              'raster')
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) install.packages(packages[!installed_packages])
lapply(packages, library, character.only = TRUE) |> invisible()
```

```{r write_bib, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::write_bib(c(.packages(), packages), 'bib/packages.bib')
```

## About 

Sometimes it is important to know the number/proportion of cells in a raster that meet a certain threshold. This can easily be done as follows:

## Creating sample raster

```{r}
set.seed(123)
r <- raster(nrow = 10, ncol = 10) # Creates the raster.
r[] <- rnorm(1:ncell(r)) # Fills the raster with random values around 0.
plot(r) # Visualizes the raster.
```

## Count number of cells in the raster

```{r}
ncell(r) # Counts the number of cells in the raster.
```

## Count cells with values above 0.3

```{r}
sum(r[] > 0.3) # Returns sum of number of cells with pixel values above 0.3.
```

## Get proportion of cells with pixel values above 0.3

```{r}
sum(r[] > 0.3) / ncell(r) # Divides sum of cells above 0.3 by total number of cells.
```

## As a percentage

```{r}
print(paste0(sum(r[] > 0.3) / ncell(r) * 100, '%')) 
```

## Application

Let's assume our raster showed suitability scores of our study area for a specific species. So, we are aware that `r sum(r[] > 0.3) / ncell(r) * 100` % of our predicted area is suitable for the species. This value is normally useful when working around significance of the built model before interpreting results.

## References





























