---
title: "<center>**... in R**</center>"
author: "<center>Wyclife Agumba Oluoch (wyclifeoluoch@gmail.com) </center>"
date: "<center>`r Sys.time()`</center>"
bibliography: 
  - bib/packages.bib
nocite: '@*'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: darkly
    #highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
packages <- c("base",
              'knitr',
              'rmarkdown',
              'prettydoc',
              'raster')
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) install.packages(packages[!installed_packages])
lapply(packages, library, character.only = TRUE) |> invisible()
```

```{r write_bib, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::write_bib(c(.packages(), packages), 'bib/packages.bib')
```

## About 

Sometimes it is important to know the number/proportion of cells in a raster that meet a certain threshold. This can easily be done as follows:

## Creating a sample raster

```{r}
set.seed(123)
r <- raster(nrow = 10, ncol = 10) # Creates the raster.
r[] <- rnorm(1:ncell(r)) # Fills the raster with random values around 0.
plot(r) # Visualizes the raster.
```

## Count number of cells in the raster

```{r}
ncell(r) # Returns the number of cells in the raster.
```

## Count cells with values > 0.3

```{r}
sum(r[] > 0.3) # Returns sum of number of cells with pixel values above 0.3.
```

## Get proportion of cells with pixel values above 0.3

```{r}
sum(r[] > 0.3) / ncell(r) # Divides sum of cells above 0.3 by total number of cells.
```

## As a percentage

```{r}
print(paste0(sum(r[] > 0.3) / ncell(r) * 100, '%')) 
```

## Application

Let's assume our raster showed suitability scores of our study area for a specific species. So, we are aware that `r sum(r[] > 0.3) / ncell(r) * 100` % of our predicted area is suitable for the species. This value is normally useful when working around significance of the built model before interpreting results.

## References

```{r}
library(tidyverse)
data.frame(x = c(31, 55, 78, 85, 94, 102,70, 65, 77.5, 54, 18, 82, 91, 53, 40, 47, 42, 57.5, 12.4, 52, 32, 80, 46, 18.9, 50, 50, 40, 102, 45, 116, 89, 60, 70, 46, 65, 60, 52, 25, 75, 25.5, 60, 35, 36, 72, 40, 52, 111, 63, 40, 75, 100, 32, 12, 36, 48, 50, 37, 17, 18, 44, 24, 24, 30, 46.6, 40, 65, 55, 48, 15, 146, 44.6, 120, 60.5, 32, 65, 50, 41, 33, 70, 21, 48.5, 37.5, 50, 29, 54, 61, 43, 42, 85, 50, 17, 33, 50, 65, 60, 70, 55, 3.5, 24, 57, 35.5, 17, 65, 76, 30, 66, 37.5, 41, 55, 48, 34, 52, 24, 60, 25, 80, 40, 36, 65, 69, 33, 11, 64, 23, 52, 35, 70.1, 100, 36, 39, 55, 80, 60, 40, 17, 41, 15, 50, 42, 37, 40, 35, 2, 125, 16, 30, 56, 15.998, 10, 40, 40, 62, 60, 43, 50, 32.5, 75, 43, 75, 30.5, 67.7, 55, 42, 50, 80, 54, 40, 20, 55)) |> ggplot(aes(x)) + geom_histogram(aes(alpha = 0.7)) + theme_classic()
```